---
output:
  revealjs::revealjs_presentation:
    theme: dark
    highlight: pygments
    center: true
    transition: fade
    reveal_options:
      slideNumber: true
      previewLinks: true
      self_contained: false
      reveal_plugins: ["notes"]
      fig_width: 10
      fig_height: 7
      fig_caption: true
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(readr)
library(tidyverse)
library(styler)
library(purrr)
library(ggthemes)
library(ggmap)
library(ggplot2)
library(ggthemes)
library(purrr)
library(ggrepel)
library(revealjs)
```
# Global Terrorism Database (GTD) using GGplot and Shiny   {data-background="http://static3.uk.businessinsider.com/image/555bc021dd08951f418b457f/world-1.gif"}
  > Authors: "Sean E. Curl & Steven Aguilar"
  > Date: "April 25, 2018"



## About
> - Global Terrorism Database (GTD)
> - Study of Terrorism and Responses to Terrorism (START)
> - University of Maryland

https://www.kaggle.com/START-UMD/gtd

##
> Definition of terrorism: 

> - "The threatened or actual use of illegal force and violence by a non-state actor to attain a political, economic, religious, or social goal through fear, coercion, or intimidation."

##

> Observations

> - The data-set included 170,350 observations.

## 

> Variables?

> - Great than 100 variables on location, tactics, perpetrators, targets, and outcomes.

## 

See the http://start.umd.edu/gtd/downloads/Codebook.pdf for important details on data collection methodology, definitions, and coding schema.


# Load and Clean {data-background="https://media.giphy.com/media/aQCCNezRpb9Hq/giphy.gif"}
##
```{r echo=FALSE, message=FALSE, warning=FALSE}
terrorism <- read_xlsx(path = "./globalterrorismdb_0617dist.xlsx", col_types = "text") %>%
  select("Year" = iyear, 
         "Month" = imonth, 
         "Day" = iday, 
         "Country" = country_txt, 
         "Region" = region_txt, 
         "AttackType" = attacktype1_txt, 
         "Target" = target1, 
         "Killed" = nkill, 
         "Wounded" = nwound, 
         "Summary" = summary, 
         "Group" = gname, 
         "Target_Type" = targtype1_txt, 
         "Weapon_type" = weapsubtype1_txt, 
         "Motive" = motive,
         "City" = city,
         "lat" = latitude,
         "long" = longitude,
         "City" = city) %>%
  
  mutate(Killed = as.numeric(Killed),
         Wounded = as.numeric(Wounded), lat = as.numeric(lat), long = as.numeric(long)) %>%
  mutate(Casualties = Killed + Wounded) %>%
  glimpse()

terrorism %>% 
  map_dbl(~sum(is.na(.)))
```


# High Impact Areas {data-background="https://media.giphy.com/media/5fGQIQ8BBhnXO/giphy.gif"}
##
```{r echo=FALSE, message=FALSE, warning=FALSE}

Year_highest_attacks <- terrorism %>%
  group_by(Year) %>%
  summarise(n = n()) %>%
  arrange(desc(n)) %>%
  slice(1)

Month_highest_attacks <- terrorism %>%
  group_by(Month) %>%
  summarise(n = mean(n())) %>%
  arrange(desc(n)) %>%
  slice(1)

Day_highest_attacks <- terrorism %>%
  group_by(Day) %>%
  summarise(n = mean(n())) %>%
  arrange(desc(n)) %>%
  slice(1)

country_highest_attacks <- terrorism %>%
  group_by(Country) %>%
  summarise(n = n()) %>%
  arrange(desc(n)) %>%
  slice(1)

region_highest_attacks <- terrorism %>%
  group_by(Region) %>%
  summarise(n = n()) %>%
  arrange(desc(n)) %>%
  slice(1)

AttackType_highest <- terrorism %>%
  group_by(AttackType) %>%
  summarise(n = n()) %>%
  arrange(desc(n)) %>%
  slice(1)

Target_highest <- terrorism %>%
  group_by(Target) %>%
  summarise(n = n()) %>%
  arrange(desc(n)) %>%
  slice(1)

highest_killed <- terrorism %>%
  group_by(Year, Country, Killed) %>%
  summarise(max(Killed)) %>%
  arrange(desc(Killed)) %>%
  head(1)

highest_wounded <- terrorism %>%
  group_by(Year, Country, Wounded) %>%
  summarise(max(Wounded)) %>%
  arrange(desc(Wounded)) %>%
  head(1)

Group_highest <- terrorism %>%
  filter(!Group == "Unknown") %>%
  group_by(Group) %>%
  summarise(n = n()) %>%
  arrange(desc(n)) %>%
  slice(1)


paste("Year with Highest Terrorist Attacks:", Year_highest_attacks$Year)
paste("Month with Highest Terrorist Attacks:", Month_highest_attacks$Month)
paste("Day with Highest Terrorist Attacks:", Day_highest_attacks$Day)
paste("Country with Highest Terrorist Attacks:", country_highest_attacks$Country)
paste("Region with Highest Terrorist Attacks:", region_highest_attacks$Region)
paste("AttackType with Highest Terrorist Attacks:", AttackType_highest$AttackType)
paste("Target with Highest Terrorist Attacks:", Target_highest$Target)
paste("Maximum peopled killed in an attack are:", "In", highest_killed$Year, highest_killed$`max(Killed)`, "peopled died in",highest_killed$Country)
paste("Maximum peopled wounded in an attack are:", "In", highest_wounded$Year, highest_wounded$`max(Wounded)`, "peopled died in",highest_wounded$Country)
paste("Group with Highest Terrorist Attacks (not unk.):", Group_highest$Group)
```
 

# Popular Attack Types {data-background="https://media.giphy.com/media/kHuMS1H30iqyc/giphy.gif"}
##
```{r echo=FALSE, message=FALSE, warning=FALSE}
attacking_methods <- terrorism %>%
  group_by(AttackType) %>%
  summarise(n = n()) %>%
  arrange(n) %>%
  ggplot(mapping = aes(reorder(AttackType, n), n, fill = AttackType)) + 
  geom_bar(stat = "identity", show.legend = FALSE) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 10)) +
  ylab("Count") +
  xlab("Attack Type") + 
  ggtitle("Attack Methods Used") +
  theme_set(theme_bw()) +
  labs(subtitle = "1970 - 2016", caption = "Source: GTD") +
  coord_flip()

attacking_methods
```


## Here's a table view. . .
```{r echo=FALSE, message=FALSE, warning=FALSE}
attacking_methods_table <- terrorism %>%
  group_by(AttackType) %>%
  summarise(n = n()) %>%
  arrange(desc(n))

print(attacking_methods_table)
```


## Let's take a look at the top three attack types. I'll include Year, Group, and Country.

##
```{r echo=FALSE, message=FALSE, warning=FALSE}
attacking_methods_table <- terrorism %>%
  filter(!Group == "Unknown") %>%
  group_by(AttackType, Year, Country, Group) %>%
  summarise(n = max(n())) %>%
  arrange(desc(n)) %>%
  head(1)

attacking_methods_table2 <- terrorism %>%
  filter(AttackType == "Armed Assault") %>%
  group_by(AttackType, Year, Country, Group) %>%
  summarise(n = max(n())) %>%
  arrange(desc(n)) %>%
  head(1)

attacking_methods_table3 <- terrorism %>%
  filter(AttackType == "Assassination") %>%
  group_by(AttackType, Year, Country, Group) %>%
  summarise(n = max(n())) %>%
  arrange(desc(n)) %>%
  head(1)

union1 <- union(attacking_methods_table, attacking_methods_table2)
union2 <- union(union1, attacking_methods_table3)
union2
```


# Number of Terrorist Acivities by Year {data-background="https://media.giphy.com/media/Cx28HgbGAwFri/giphy.gif"}
##
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Number of Terrorist Acivities Year
count <- terrorism %>%
  group_by(Year) %>%
  summarise(n = n()) %>%
  mutate(Year = as.numeric(Year))

ggplot(count, 
       aes(x = Year, 
           y = n, fill = factor(Year))) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  xlab("Year") +
  ylab("Count") +
  ggtitle("Global Number of Terrorist Acivities") +
  scale_x_continuous(breaks = c(1970, 1980, 1990, 2000, 2010, 2020), limits = c(1970, 2016)) +
  theme_set(theme_bw()) +
  labs(subtitle = "1970 - 2016",
       caption = "Source: GTD") +
  coord_flip()
```


##  We can see that the most recent years had a significant increase in violence.

##
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Number of Terrorist Acivities, Top Six Years
count2 <- terrorism %>%
  group_by(Year) %>%
   summarise(n = n()) %>%
  arrange(desc(n))

print(head(count2))
```


## Let's take a look at the Year's > 2010.

##
```{r echo=FALSE, message=FALSE, warning=FALSE}
Year_Great_2010 <- terrorism %>%
  filter(!Group == "Unknown") %>%
  filter(Year >= '2010' & Year <= '2016') %>%
  group_by(Year, Group) %>%
  summarise(n = max(n())) %>%
  arrange(desc(n))
  
output1 <- distinct(Year_Great_2010, Country, .keep_all = TRUE)
head(output1, 5)
```


## How about looking at the Year's < 2010?

##
```{r echo=FALSE, message=FALSE, warning=FALSE}
Year_Less_2010 <- terrorism %>%
  filter(!Group == "Unknown") %>%
  filter(Year < '2010') %>%
  group_by(Year, Group) %>%
  summarise(n = max(n())) %>%
  arrange(desc(n))

output <- distinct(Year_Less_2010, Year, .keep_all = TRUE)
head(output, 5)
```


# Terrorist Targets by Count {data-background="https://media.giphy.com/media/NTkXyye1ejDIk/giphy.gif"}
##
```{r echo=FALSE, message=FALSE, warning=FALSE}
count <- terrorism %>%
  group_by(Target_Type) %>%
  summarise(n = n()) %>%
  head(10)

ggplot(count, 
       aes(reorder(x = Target_Type, n), 
           y = n, fill = factor(Target_Type))) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  xlab("Target Type") +
  ylab("Count") +
  ggtitle("Global Terrorist Targets") +
  theme(axis.text.x=element_text(angle = 90, hjust =1)) +
  theme_set(theme_bw()) +
  labs(subtitle = "by Count",
       caption = "Source: GTD") +
  coord_flip()
```


## What are the top 5 preferred targets for different groups?

##
```{r echo=FALSE, message=FALSE, warning=FALSE}
Target_Group <- terrorism %>%
  filter(!Group == "Unknown") %>%
  group_by(Target_Type, Group) %>%
  summarise(n = max(n())) %>%
  arrange(desc(n))

output <- distinct(Target_Group, Target_Type, .keep_all = TRUE)
head(output, 5)
```


## What are top 5 preferred targets within different countries?

##
```{r echo=FALSE, message=FALSE, warning=FALSE}
Target_Country <- terrorism %>%
  group_by(Target_Type, Country) %>%
  summarise(n = max(n())) %>%
  arrange(desc(n))

output <- distinct(Target_Country, Target_Type, .keep_all = TRUE)
head(output, 5)
```

# Terrorist Attacks by City in USA {data-background="https://media.giphy.com/media/xUNd9FaInWieZLffig/giphy.gif"}

##
```{r echo=FALSE, message=FALSE, warning=FALSE}
highest_USA <- terrorism %>%
    filter(City %in% c("New York City", "Oklahoma City",
                                            "The Dalles", 
                                            "Arlington", 
                                            "West",
                                            "Boston",
                                            "Atlanta",
                                            "Orlando")) %>%
  group_by(City) %>%
  summarise(total = sum(n())) %>%
  arrange(desc(total))

ggplot(highest_USA, 
       aes(reorder(x = City, total), 
           y = total, fill = factor(City))) +
  geom_bar(stat = "identity", width = .5, fill = "red", show.legend = FALSE) +
  xlab("City Name") +
  ylab("Count") +
    ggtitle("Highest Number of U.S. City Terrorist Attacks") +
  labs(subtitle = "with New York City",
       caption = "Source: GTD") +
  theme_set(theme_bw()) +
  theme(axis.text.x = element_text(vjust=0.6)) +
  coord_flip()

```

## New York City has the highest number of terrorist attacks since 1970.

##
```{r echo=FALSE, message=FALSE, warning=FALSE}
count <- terrorism %>%
  filter(City == "New York City") %>%
  group_by(City, Year) %>%
  summarise(total = sum(n())) %>%
  arrange(desc(total))

output <- distinct(count, Year, .keep_all = TRUE)
head(output, 10)
```

## What do these attacks look like across time?

##
```{r echo=FALSE, message=FALSE, warning=FALSE}
highest_USA <- terrorism %>%
    filter(City %in% c("New York City", "Oklahoma City",
                                            "The Dalles", 
                                            "Arlington", 
                                            "West",
                                            "Boston",
                                            "Atlanta",
                                            "Orlando")) %>%
  group_by(City, Year) %>%
  summarise(total = sum(n())) %>%
  arrange(desc(total)) %>%
  mutate(Year = as.numeric(Year))


ggplot(highest_USA, 
       aes(x = Year, 
           y = total, colour = City, group = City)) +
  geom_line(stat = "identity", lwd = 1) +
  xlab("Year") +
  ylab("Count") +
  scale_x_continuous(breaks = c(1970, 1980, 1990, 2000, 2010, 2020), limits = c(1970, 2016)) +
  ggtitle("U.S. City Terrorist Attacks") +
  theme(legend.title = element_text(color = "black", size = 10, face = "bold")) + 
  theme_set(theme_bw()) +
  labs(subtitle = "by City",
       caption = "Source: GTD")
```

## Across all U.S. cities, which had the deadliest attacks?

##
```{r echo=FALSE, message=FALSE, warning=FALSE}
highest_killed_USA <- terrorism %>%
  filter(Country == "United States")  %>%
  group_by(Year, City, Casualties) %>%
  summarise(total = sum(Casualties)) %>%
  arrange(desc(total)) 

output <- distinct(highest_killed_USA, City, .keep_all = TRUE)
head(output, 10)
```

## Let's exclude New York City.

## 
```{r echo=FALSE, message=FALSE, warning=FALSE}
count <- terrorism %>%
    filter(City %in% c("Oklahoma City", "Hyder", "Shanksville",
                                            "The Dalles", 
                                            "Arlington", 
                                            "West",
                                            "Boston",
                                            "Atlanta",
                                            "Orlando")) %>%
  group_by(City) %>%
  summarise(total = sum(n())) %>%
  arrange(desc(total))

ggplot(count,
       aes(reorder(x = City, total), 
           y = total, fill = factor(City))) +
  geom_bar(stat = "identity", width = .5, fill = "red", show.legend = FALSE) +
  xlab("City Name") +
  ylab("Count") +
  ggtitle("Terrorist Attacks in Top U.S. Cities") +
  theme_set(theme_bw()) +
  labs(subtitle = "Minus New York City",
       caption = "Source: GTD") +
  coord_flip()
```

##
```{r echo=FALSE, message=FALSE, warning=FALSE}
count <- terrorism %>%
    filter(City %in% c("Oklahoma City",
                                            "The Dalles", 
                                            "Arlington", 
                                            "West",
                                            "Boston",
                                            "Atlanta",
                                            "Orlando")) %>%
  group_by(City, Year) %>%
  summarise(total = sum(n())) %>%
  arrange(desc(total))

output <- distinct(count, City, .keep_all = TRUE)
head(output)
```

## Again, what do these attacks look like across time?

##
```{r echo=FALSE, message=FALSE, warning=FALSE}
highest_USA <- terrorism %>%
    filter(City %in% c("Oklahoma City",
                                            "The Dalles", 
                                            "Arlington", 
                                            "West",
                                            "Boston",
                                            "Atlanta",
                                            "Orlando")) %>%
  group_by(City, Year) %>%
  summarise(total = sum(n())) %>%
  arrange(desc(total)) %>%
  mutate(Year = as.numeric(Year))


ggplot(highest_USA, 
       aes(x = Year, 
           y = total, colour = City, group = City)) +
  geom_bar(stat = "identity", lwd = 1) +
  xlab("Year") +
  ylab("Count") +
  scale_x_continuous(breaks = c(1970, 1980, 1990, 2000, 2010, 2020), limits = c(1970, 2016)) +
  ggtitle("U.S. City Terrorist Attacks") +
  theme(legend.title = element_text(color = "black", size = 10, face = "bold")) + 
  theme_set(theme_bw()) +
  labs(subtitle = "by City (w/o New York City)",
       caption = "Source: GTD")
```

## Across these U.S. cities, which had the deadliest attacks?

##
```{r echo=FALSE, message=FALSE, warning=FALSE}
highest_killed_USA <- terrorism %>%
  filter(Country == "United States")  %>%
  filter(!City == "New York City")  %>%
  group_by(Year, City, Casualties) %>%
  summarise(total = sum(Casualties)) %>%
  arrange(desc(total))

output <- distinct(highest_killed_USA, Year, .keep_all = TRUE)
head(output)
```

## {data-background-iframe="https://saguilar.shinyapps.io/GTDB/"}


## Presentation created using REVEAL.JS
> - https://revealjs.com/#/
> - https://rmarkdown.rstudio.com/revealjs_presentation_format.html
> - http://rpubs.com/securl/revealjspresent2

## Github Repo:
> - https://github.com/dataSeanC

## QUESTIONS? {data-background="https://media.giphy.com/media/xUNd9FaInWieZLffig/giphy.gif"}
